Theory about Normal Basis on $F(2^k)$

### 2017-03-03 dissertation/submission/prelim_basis.tex
after submission: revision 2

\chapter{Gr\"ober Bases and Algebraic Geometry} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly referred from the textbooks \cite{ideals:book,gb_book} and 
previous work by Lv \cite{lv:phd} as well as Pruss \cite{pruss:tcad15}.

@@

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, {\it i.e.} $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.

@@

A {\bf degree-lexicographic ordering} (DEGLEX) is a total-ordering $>$ such 
that the total degree of a term takes precedence over the lexicographic 
ordering.  
A {\bf degree-reverse-lexicographic ordering} (DEGREVLEX) is the same as a
DEGLEX ordering, however terms are LEXed in reverse.

@@

\begin{Definition}
Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
\begin{equation}
f-\frac{lt(f)}{lt(g)}\cdot g
\end{equation}
This polynomial division is denoted
\begin{equation}
f\xrightarrow{g} r
\end{equation}
where $r$ is the resulting polynomial of the division.
If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
{\it i.e.} $g \mid f$.
\end{Definition}
Notice that if $g \nmid f$, that is if $f$ is not divisible by $g$,
then the division operation gives $r = f$.
\begin{Example}
Over $\R[x,y,z]$, set the lex term order $x > y > z$.
Let $f = -2x^3 + 2x^2yz + 3xy^3$ and $g = x^2+yz$.
\begin{equation}
\frac{lt(f)}{lt(g)} = \frac{-2x^3}{x^2} = -2x
\end{equation}
Since $\frac{lt(f)}{lt(g)}$ is non-zero $g|f$. The division, $f\xrightarrow{g} r$, 
is computed as:
\begin{eqnarray}
r &=& f-\frac{lt(f)}{lt(g)}\cdot g = -2x^3 + 2x^2yz + 3xy^3 - (-2x \cdot (x^2+yz)) \nonumber \\
&=& -2x^3 + 2x^2yz + 3xy^3 - (-2x^3-2xyz) = 2x^2yz + 3xy^3 + 2xyz
\end{eqnarray}
Notice that the division $f\stackrel{g}{\textstyle\longrightarrow}r$ cancels the leading term of $f$.
\end{Example}

@@

    For sets of polynomials, the notation 
    \begin{equation*}
    f\stackrel{F}{\textstyle\longrightarrow}_+r    
    \end{equation*}
    represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
    set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
    $r = 0$  or no term in $r$ is divisible  by $lm(f_i), \forall f_i \in F$.
\end{Definition}

@@

\subsection{Varieties and Ideals}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In computer-algebra based formal verification,
it is often necessary to analyze the presence 
or absence of solutions to a given system of constraints.
In our applications, these constraints are polynomials and their solutions 
are modeled as {\bf varieties}.

@@

The above example shows the variety can be infinite, finite (non-empty set) 
or empty. It is interesting to note that since we will be operating over
finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
any variety over $\F_{q}$ is finite (or empty).
Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
in $\F_q^d$. Any single point is a variety of some polynomial system:
{\it e.g.} $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
varieties are also varieties.

@@

The same ideal may have many different bases.
For instance, it is possible to have different sets of polynomials
$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
ideal, {\it i.e.} 
$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
variety depends on the ideal, these sets of polynomials have the same 
solutions.

@@

Ideals and their varieties are a key part of computer-algebra based formal
verification. A given hardware design can be transformed into a set
of polynomials over a field, $F = \{f_1, \ldots, f_s\}$.
This set of polynomials gives the system of equations:
\begin{eqnarray}
f_1 = 0 \nonumber \\
\vdots \nonumber \\
f_s = 0 \nonumber
\end{eqnarray}
Using algebra, it is possible to derive new equations from the original 
system.
The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
such {\it consequences} of a system of polynomials.

\begin{Example}
Given two equations in $\mathbb{R}[x,y,z]$:
\begin{eqnarray}
x=z+1 \nonumber \\
y=x^2+1 \nonumber 
\end{eqnarray}
we can eliminate $x$ to obtain a new equation:
\begin{equation}
y=(z+1)^2+1=z^2+2z+2 \nonumber 
\end{equation}
Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
equations:
\begin{eqnarray}
f_1 &= x-z-1 &= 0 \nonumber \\
f_2 &= y-x^2-1 &= 0 \nonumber \\
h   &= y-z^2-2z-2 &= 0 \nonumber
\end{eqnarray}
If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.} 
$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
\begin{eqnarray}
g_1 &= x+z+1 \nonumber \\
g_2 &= 1     \nonumber \\
h &= g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
\end{eqnarray}
where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
Thus, we call $h$ a {\bf member of the ideal} $I$.
\end{Example}

@@

\Grobner basis has an important property, and therefore can be used to perform ideal membership test.
Formally speaking, a \Grobner basis gives a decision procedure to test 
for polynomial membership in an ideal. This is explained in the following 
Theorem.

@@

\begin{Example}
Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
by $f_1$ first and then by $f_2$, we will obtain a zero
remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
basis, we find that the reduction $f
\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
{\it i.e.} dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
in $G$ in any order will always lead to the zero remainder. Therefore,
one can decide ideal membership unequivocally using the Gr\"{o}bner
basis. 
\end{Example}

@@

Buchberger's algorithm takes pairs of polynomials ($f_{i}, f_{j}$) in 
the basis $G$ and combines them into ``$S$-polynomials'' 
($Spoly(f_{i}, f_{j})$) to cancel leading terms. The $S$-polynomial is then 
reduced (divided) by all elements of $G$ to a remainder $r$, denoted as  
$Spoly(f_{i}, f_{j}) \stackrel{G}{\textstyle\longrightarrow}_+r$. This
process is repeated for all unique pairs of polynomials, including
those created by newly added elements, until no new polynomials are
generated; ultimately constructing the \Grobner basis.
\begin{Example}\label{exp:gbsimple}
Consider the ideal $I \subset \mathbb{Q}[x, y]$, $I = \langle f_1, f_2 
\rangle$, where $f_1 = yx - y, ~f_2 = y^2 - x$. 
Assume a degree-lexicographic term ordering with $y > x$ is imposed.

@@

When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
and $lt(f_j)$ have no common variables, the $Spoly$ reduction step in 
Buchberger's algorithm will reduce to 0.
\begin{Lemma}
In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the $Spoly$ reduction
$$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
will produce $r=0$.
\end{Lemma}

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
{\it i.e.} remove every polynomial in $G$ whose leading term is divisible by the leading term 
of some other polynomial in $G$.
\end{enumerate}

@@

The high computational complexity of the Gr\"obner basis problem is an important issue
because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential, as the total degree of 
polynomials in the Gr\"obner basis is bounded by 
$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretical basis of this dissertation.

@@

%%%%%%%%%%%%%%%%%algebraically closed field%%%%%%%%%%%%%
\begin{Definition}\label{def:acf}
A field $\overline {\F}$ is an {\bf algebraically closed} field if every  
polynomial in one variable with degree at least $1$, with coefficients 
in $\overline {\F}$, has a root in $\overline {\F}$. 
\end{Definition}
In other words, any non-constant polynomial equation over 
$\overline {\F}\left[x\right]$ always has at least one root 
in $\overline {\F}$. Every field $\F$ is contained in an algebraically 
closed one $\overline {\F}$. 
For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
field, because $x^2+1=0$ has no root in $\mathbb{R}$. 
However, $x^2+1=0$ has roots in the field of 
complex numbers $\mathbb{C}$, which is an algebraically closed field. 
In fact, $\mathbb{C}$ is the algebra closure of $\mathbb{R}$. 
Every algebraically closed field is an infinite field.

%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
\begin{Theorem}[Weak Nullstellensatz]
%$\left[\bf{Weak\  Nullstellensatz}\right]$ 
Let $J \subset \overline {\F}[x_1, x_2, \cdots, x_d]$ 
be an ideal satisfying $V(J)=\emptyset$. 
Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
\begin{equation}
V(J)=\emptyset\ \iff\ J=\overline {\F}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
\end{equation}
\end{Theorem}

@@

\begin{Theorem}[Strong Nullstellensatz]
\label{thm:sns}
\cite{gb_book} 
Let $\overline{\F}$ be an algebraically closed field, and let $J$
be an ideal in $\overline{\F}[x_1,\dots, x_d]$. 
Then we have $I(V_{\overline{\F}}(J)) =\sqrt{J}$. 
\end{Theorem}

@@

Using this result, Weak Nullstellensatz can be 
modified to be applicable over finite fields $\Fq$.
%%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
\begin{Theorem}[Weak Nullstellensatz in $\Fq$]
\label{wnull:ff}
\cite{null:1890}
Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
of vanishing polynomials in $\Fq$. Then
$V_{\Fq}(J) = V_{\overline {\Fq}}(J +
J_0)=\emptyset$,  if and only if the reduced
Gr\"obner Basis $RGB(J+J_{0})=\{1\}$. 
\end{Theorem}

@@

\section{Elimination Theory and Application to Abstraction}
\label{sec:abstraction}
Elimination of certain variables in a system of polynomials is a common 
operation when some variables is not needed in modeling and analysis. 
In this section, eliminating variables targets a tight-bound over-approximation.
This is equivalent to the existential quantifier elimination in first-order logic,
or variable smoothing in Boolean operations. We introduce a elimination method 
based on algebraic geometry concepts, and use it as the fundamental of 
abstraction of a circuit.

@@

Formally, we define concepts of projection and elimination ideal:
\begin{Definition}
The {\bf $l$-th projection} mapping is defined as:
$$\pi_l:\Fq^d\to\Fq^{d-l},~~\pi_l((x_1,\dots,x_d)) = (x_{l+1}.\dots,x_d)$$
where $l<d$. For any subset $V \subseteq \Fq^d$, we write 
$$\pi_l(V) = \{\pi_i(x):x\in A\}\subseteq \Fq^{d-l}$$
\end{Definition}

@@

The elimination ideal $J_l$ has eliminated all the variables 
$x_1,\dots,x_l$, {\it i.e.} it only contains polynomials with variables in
$x_{l+1},\dots,x_d$. 
In a general setting,
$\pi_l(V_d(J))$ from $\Fq^d$ to $\Fq^{d-l}$ is a subset of $V(J_l)$. But over $\Fq$, 
consider ideal $J+J_0$ where $J_0$ is the ideal of vanishing polynomials, it is proved in \cite{gao:qe-gf-gb}
that 
$$\pi_l(V(J+J_0)) = V((J+J_0)_l)$$

This shows that in finite fields, projection of the variety $V_{d}(\langle f_1,\dots,f_s\rangle)$
from $\Fq^d$ to $\Fq^{d-l}$, is exactly the variety $V_{d-l}(\langle f_1,\dots,f_s\rangle\cap \Fq[x_{l+1},\dots,x_d])$.

@@

Elimination theory provides the basis for following abstraction approach.
\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
\begin{Problem}
\label{setup:abs}
Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of any circuit over $\Fq, q=2^k$, as Figure \ref{fig:abs_setup} shows.
In general, this circuit performs some unknown function $f$ over 
$\Fq$ such that every $f:\Fq\to\Fq$ is represented in polynomial form.
For any circuit $\mathcal C$, $Z = \Func(A)$. Then how to derive this function $\Func$
using GB?
\end{Problem}

@@

Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
it agrees with the solution to the system of polynomials $\{S\}=0$, {\it i.e.}:
\begin{equation}
f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
\end{equation}
Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
be the ideal generated by $S$, 
$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
Therefore, due to Proposition \ref{pro:iofv}, 
$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.

>>>>

\chapter{Gr\"ober Bases and Algebraic Geometry} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly taken from the textbooks \cite{ideals:book,gb_book} and 
previous work by Lv \cite{lv:phd} as well as Pruss \cite{pruss:tcad15}.

@@

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.

@@

A {\bf degree-lexicographic ordering} (DEGLEX) is a total-ordering $>$ such 
that the total degree of a term takes precedence over the lexicographic 
ordering.  
A {\bf degree-reverse-lexicographic ordering} (DEGREVLEX) is the same as a
DEGLEX ordering, but terms are LEXed in reverse.

@@

\begin{Definition}
Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
over the field. {\bf Polynomial division} of $f$ by $g$ is computed following:
\begin{equation}
f-\frac{lt(f)}{lt(g)}\cdot g
\end{equation}
This polynomial division is denoted
\begin{equation}
f\xrightarrow{g} r
\end{equation}
where $r$ is the resulting polynomial of the division.
If $\frac{lt(f)}{lt(g)}$ is nonzero, then $f$ is considered divisible by $g$, 
{\it i.e.}, $g \mid f$.
\end{Definition}
Notice that if $g \nmid f$, that is if $f$ is not divisible by $g$,
then the division operation gives $r = f$.
\begin{Example}
Over $\R[x,y,z]$, set the lex term order $x > y > z$.
Let $f = -2x^3 + 2x^2yz + 3xy^3$ and $g = x^2+yz$.
\begin{equation}
\frac{lt(f)}{lt(g)} = \frac{-2x^3}{x^2} = -2x
\end{equation}
Since $\frac{lt(f)}{lt(g)}$ is nonzero $g|f$. The division, $f\xrightarrow{g} r$, 
is computed as:
\begin{eqnarray}
r &=& f-\frac{lt(f)}{lt(g)}\cdot g = -2x^3 + 2x^2yz + 3xy^3 - (-2x \cdot (x^2+yz)) \nonumber \\
&=& -2x^3 + 2x^2yz + 3xy^3 - (-2x^3-2xyz) = 2x^2yz + 3xy^3 + 2xyz
\end{eqnarray}
Notice that the division $f\stackrel{g}{\textstyle\longrightarrow}r$ cancels the leading term of $f$.
\end{Example}

@@

    For sets of polynomials, the notation 
    \begin{equation*}
    f\stackrel{F}{\textstyle\longrightarrow}_+r    
    \end{equation*}
    represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
    set of nonzero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
    $r = 0$  or no term in $r$ is divisible  by $lm(f_i), \forall f_i \in F$.
\end{Definition}

@@

\subsection{Varieties and Ideals}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In computer-algebra-based formal verification,
it is often necessary to analyze the presence 
or absence of solutions to a given system of constraints.
In our applications, these constraints are polynomials and their solutions 
are modeled as {\bf varieties}.

@@

The above example shows the variety can be infinite, finite (nonempty set) 
or empty. It is interesting to note that since we will be operating over
finite fields $\F_{q}$, any finite set of points is a variety. Likewise,
any variety over $\F_{q}$ is finite (or empty).
Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
in $\F_q^d$. Any single point is a variety of some polynomial system:
{\it e.g.} $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
varieties are also varieties.

@@

The same ideal may have many different bases.
For instance, it is possible to have different sets of polynomials
$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
ideal, i.e., $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
variety depends on the ideal, these sets of polynomials have the same 
solutions.

@@

Ideals and their varieties are a key part of computer-algebra-based formal
verification. A given hardware design can be transformed into a set
of polynomials over a field, $F = \{f_1, \ldots, f_s\}$.
This set of polynomials gives the system of equations:
\begin{eqnarray}
f_1 = 0 \nonumber \\
\vdots \nonumber \\
f_s = 0 \nonumber
\end{eqnarray}
Using algebra, it is possible to derive new equations from the original 
system.
The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
such {\it consequences} of a system of polynomials.

\begin{Example}
Given two equations in $\mathbb{R}[x,y,z]$:
\begin{eqnarray}
x=z+1 \nonumber \\
y=x^2+1 \nonumber 
\end{eqnarray}
we can eliminate $x$ to obtain a new equation:
\begin{equation}
y=(z+1)^2+1=z^2+2z+2 \nonumber 
\end{equation}
Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
equations:
\begin{eqnarray}
f_1 &= x-z-1 &= 0 \nonumber \\
f_2 &= y-x^2-1 &= 0 \nonumber \\
h   &= y-z^2-2z-2 &= 0 \nonumber
\end{eqnarray}
If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.}, 
$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
\begin{eqnarray}
g_1 &= x+z+1 \nonumber \\
g_2 &= 1     \nonumber \\
h &= g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
\end{eqnarray}
where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
Thus, we call $h$ a {\bf member of the ideal} $I$.
\end{Example}

@@

\Grobner basis has an important property, and therefore can be used to perform an ideal membership test.
Formally speaking, a \Grobner basis gives a decision procedure to test 
for polynomial membership in an ideal. This is explained in the following 
Theorem.

@@

\begin{Example}
Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
by $f_1$ first and then by $f_2$, we will obtain a zero
remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
basis, we find that the reduction $f
\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
{\it i.e.}, dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
in $G$ in any order will always lead to the zero remainder. Therefore,
one can decide ideal membership unequivocally using the Gr\"{o}bner
basis. 
\end{Example}

@@

Buchberger's algorithm takes pairs of polynomials ($f_{i}, f_{j}$) in 
the basis $G$ and combines them into ``$S$-polynomials'' 
($Spoly(f_{i}, f_{j})$) to cancel leading terms. The $S$-polynomial is then 
reduced (divided) by all elements of $G$ to a remainder $r$, denoted as  
$Spoly(f_{i}, f_{j}) \stackrel{G}{\textstyle\longrightarrow}_+r$. This
process is repeated for all unique pairs of polynomials, including
those created by newly added elements, until no new polynomials are
generated; ultimately constructing the \Grobner basis.
\begin{Example}\label{exp:gbsimple}
Consider the ideal $I \subset \mathbb{Q}[x, y]$, $I = \langle f_1, f_2 
\rangle$, where $f_1 = yx - y, ~f_2 = y^2 - x$. 
Assume a degree-lexicographic term ordering with $y > x$ imposed.

@@

When computing a \Grobner basis, it is important to note that if $lt(f_i)$ 
and $lt(f_j)$ have no common variables, the $Spoly$ reduction step in 
Buchberger's algorithm will reduce to 0.
\begin{Lemma}
In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the $Spoly$ reduction
$$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
will produce $r=0$.
\end{Lemma}

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
i.e., remove every polynomial in $G$ whose leading term is divisible by the leading term 
of some other polynomial in $G$.
\end{enumerate}

@@

The high computational complexity of the Gr\"obner basis problem is an important issue
because of its high cost in time and space. Concretely, for an arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential, as the total degree of 
polynomials in the Gr\"obner basis is bounded by 
$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
\cite{dube1986complexity}. However, in situations such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao {\it et al.} \cite{gao:qe-gf-gb} prove that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretical basis of this dissertation.

@@

%%%%%%%%%%%%%%%%%algebraically closed field%%%%%%%%%%%%%
\begin{Definition}\label{def:acf}
A field $\overline {\F}$ is an {\bf algebraically closed} field if every  
polynomial in one variable with degree at least $1$, with coefficients 
in $\overline {\F}$, has a root in $\overline {\F}$. 
\end{Definition}
In other words, any nonconstant polynomial equation over 
$\overline {\F}\left[x\right]$ always has at least one root 
in $\overline {\F}$. Every field $\F$ is contained in an algebraically 
closed one $\overline {\F}$. 
For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
field, because $x^2+1=0$ has no root in $\mathbb{R}$. 
However, $x^2+1=0$ has roots in the field of 
complex numbers $\mathbb{C}$, which is an algebraically closed field. 
In fact, $\mathbb{C}$ is the algebra closure of $\mathbb{R}$. 
Every algebraically closed field is an infinite field.

%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
\begin{Theorem}[Weak Nullstellensatz]
%$\left[\bf{Weak\  Nullstellensatz}\right]$ 
Let $J \subset \overline {\F}[x_1, x_2, \cdots, x_d]$ 
be an ideal satisfying $V(J)=\emptyset$. 
Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, or equivalently, 
\begin{equation}
V(J)=\emptyset\ \iff\ J=\overline {\F}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
\end{equation}
\end{Theorem}

@@

\begin{Theorem}[Strong Nullstellensatz \cite{gb_book}]
\label{thm:sns}
Let $\overline{\F}$ be an algebraically closed field, and let $J$
be an ideal in $\overline{\F}[x_1,\dots, x_d]$. 
Then we have $I(V_{\overline{\F}}(J)) =\sqrt{J}$. 
\end{Theorem}

@@

Using this result, Weak Nullstellensatz can be 
modified to be applicable over finite fields $\Fq$.
%%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
\begin{Theorem}[Weak Nullstellensatz in $\Fq$ \cite{null:1890}]
\label{wnull:ff}
Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
of vanishing polynomials in $\Fq$. Then
$V_{\Fq}(J) = V_{\overline {\Fq}}(J +
J_0)=\emptyset$,  if and only if the reduced
Gr\"obner Basis $RGB(J+J_{0})=\{1\}$. 
\end{Theorem}

@@

\section{Elimination Theory and Application to Abstraction}
\label{sec:abstraction}
Elimination of certain variables in a system of polynomials is a common 
operation when some variables are not needed in modeling and analysis. 
In this section, eliminating variables targets a tight-bound overapproximation.
This is equivalent to the existential quantifier elimination in first-order logic,
or variable smoothing in Boolean operations. We introduce an elimination method 
based on algebraic geometry concepts, and use it as the fundamental of 
abstraction of a circuit.

@@

Formally, we define concepts of the projection and elimination ideals:
\begin{Definition}
The {\bf $l$-th projection} mapping is defined as:
$$\pi_l:\Fq^d\to\Fq^{d-l},~~\pi_l((x_1,\dots,x_d)) = (x_{l+1}.\dots,x_d)$$
where $l<d$. For any subset $V \subseteq \Fq^d$, we write 
$$\pi_l(V) = \{\pi_i(x):x\in A\}\subseteq \Fq^{d-l}$$
\end{Definition}

@@

The elimination ideal $J_l$ has eliminated all the variables 
$x_1,\dots,x_l$, i.e., it only contains polynomials with variables in
$x_{l+1},\dots,x_d$. 
In a general setting,
$\pi_l(V_d(J))$ from $\Fq^d$ to $\Fq^{d-l}$ is a subset of $V(J_l)$. But over $\Fq$, 
consider ideal $J+J_0$ where $J_0$ is the ideal of vanishing polynomials, it is proved in \cite{gao:qe-gf-gb}
that 
$$\pi_l(V(J+J_0)) = V((J+J_0)_l).$$
This shows that in finite fields, projection of the variety $V_{d}(\langle f_1,\dots,f_s\rangle)$
from $\Fq^d$ to $\Fq^{d-l}$, is exactly the variety $V_{d-l}(\langle f_1,\dots,f_s\rangle\cap \Fq[x_{l+1},\dots,x_d])$.

@@

Elimination theory provides the basis for the following abstraction approach.
\subsection{Abstraction Using Nullstellensatz and Gr\"obner Basis}
\begin{Problem}
\label{setup:abs}
Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of any circuit over $\Fq, q=2^k$, as Figure \ref{fig:abs_setup} shows.
In general, this circuit performs some unknown function $f$ over 
$\Fq$ such that every $f:\Fq\to\Fq$ is represented in polynomial form.
For any circuit $\mathcal C$, $Z = \Func(A)$. Then how to derive this function $\Func$
using GB?
\end{Problem}

@@

Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.,:
\begin{equation}
f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
\end{equation}
Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
be the ideal generated by $S$, 
$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
Therefore, due to Proposition \ref{pro:iofv}, 
$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.
